---
title: "Text Mining - Harry Potter and the Prisoner of Azkaban"
output: html_document
date: "2025-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Potential Problems with the Script and Solutions**

**Problem:**

Mixed Narrative and Dialogue - The script contains scene descriptions (e.g., "The street slumbers") alongside dialogue (e.g., "Lumos Maxima"), which could complicate isolating dialogue or sentiment.

**Solution:** Use text patterns (e.g., character names in all caps followed by lines) to separate dialogue from narrative.

**Problem:** Limited Context - Scripts are shorter and less descriptive than books, potentially skewing sentiment or n-gram results due to fewer words.

**Solution:** Normalize results (e.g., frequencies per 1000 words) to account for length differences.

**Problem:** Stage Directions - Actions like "The tent BLOOMS SOFTLY" may interfere with word frequency or sentiment analysis if not filtered out.

**Solution:** Preprocess to exclude text within parentheses or italicized directions (if formatted consistently).

**Problem:** Character Identification - While your script labels speakers (e.g., "HARRY"), mentions of characters within dialogue or narrative may be harder to extract accurately.

**Solution:** Use named entity recognition (NER) or manual tagging to identify character mentions.

## Libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(tm)
library(textstem)
library(udpipe)
library(stopwords)
library(tidytext)
library(ggplot2)
library(syuzhet)

```


## Reading the book and script

```{r}
azkaban_book <- read_file("data/azkaban_book.txt")

azkaban_script <- read_file("data/azkaban_script.txt")

# making the book a tibble
azkaban_book <- tibble(text = azkaban_book)

# split the script into lines and store as a tibble
azkaban_script <- tibble(text = str_split(azkaban_script, pattern = "\n") |> unlist())

```

**For a comprehensive stopword removing**

```{r message=FALSE, warning=FALSE}
ud_model <- udpipe_download_model(language = "english")
ud_model <- udpipe_load_model(ud_model$file_model)
```


## Word Frequencies for Book and Script

Tokenizing splits text into individual words or units, like turning "Harry runs fast" into "Harry", "runs", "fast". 

Lemmatizing reduces words to their base form, so "runs" and "running" both become "run". 

Removing stopwords filters out common words that don’t carry much meaning, like taking "the" and "is" out of "the spell is unmerciful" to leave "spell" and "unmerciful".


### The book

'udpipe_annotate' does tokenization + POS tagging + lemmatization.

Part-of-speech (POS) tags: Identifying whether a word is a noun, verb, adjective, etc.

Lemmatization: Converting words to their base form (e.g., "running" → "run").


```{r}

# annotating text with POS tags and converting to dataframe
tokens_azkaban_book <- udpipe_annotate(ud_model, x = azkaban_book$text) |> 
  as.data.frame()

# loading built-in stopwords list
stop_words_azkaban <- stopwords(source = "smart")

# clean tokens by filtering out verbs, punctuation, pronouns, stopwords, and empty lemmas
clean_tokens_azkaban_book <- tokens_azkaban_book |>
  filter(upos != "VERB",  
         upos != "PUNCT",  
         upos != "PRON",   
         !grepl("[[:punct:]]", lemma),  
         !lemma %in% stop_words_azkaban,  
         lemma != "") |>  
  pull(lemma)  

```


#### Term Frequencies

An interesting finding from at first glance is that in the book Hagrid is mentioned more than Professor Snape and Sirius Black (However it is worth noting that Sirius Black is referred to as 'Black' and 'Sirius', which might mean that his name is mentioned more than both Hagrid and Snape). This is interesting to those who are familiar with the movie. Professor Lupin, the lycanthrope, is central to the story. 

In the script, Hagrid is given less mentions compared to Snape and Sirius Black.

```{r}

# to see clearly the most frequent words
clean_tokens_azkaban_book |> 
  table() |>  
  sort(decreasing = TRUE) |>  
  head(50)  


```


#### Term Frequencies - but as percentages

One thing we'd like to know is whether some characters are mentioned less in the script compared to the book. 

The book puts Harry, Ron, Hermione, Lupin, Hagrid, and Sirius Black in the centre of the story in Azkaban. Professor Snape is also mentioned with a fair frequency, close to that of Hagrid. 

```{r}
# get word frequency table
book_word_freq <- table(clean_tokens_azkaban_book)

# convert to percentage
word_freq_percent <- prop.table(book_word_freq) * 100

# show top 50 most frequent words with percentages
sort(word_freq_percent, decreasing = TRUE) |> head(50)

```



### The script

The script, on the other hand, keeps Harry, Ron, and Hermione central to the story, but mentioning Hermione more than Ron this time. It also mentions the word 'black', which is the last name of Sirius, more frequent than the book. However, the word 'black' can be used in other contexts as well. Hagrid falls behind Snape in the frequency of mentions in the movie script, showing that the movie gives more emphasis to Snape than Hagrid. Last but not least, Lupin becomes even more important for the story with a greater mention frequency (from 1.2% in the book to 1.57% (Lupin + LUPIN combined). 

```{r}

# annotate the script text with udpipe and convert to dataframe
tokens_azkaban_script <- udpipe_annotate(ud_model, x = azkaban_script$text) |> 
  as.data.frame()


# getting the stopwords
stop_words_script <- stopwords(source = "smart")


# clean tokens by filtering out verbs, punctuation, pronouns, stopwords, and empty lemmas
clean_tokens_azkaban_script <- tokens_azkaban_script |>
  filter(upos != "VERB",  
         upos != "PUNCT",  
         upos != "PRON",   
         !grepl("[[:punct:]]", lemma),  
         !lemma %in% stop_words_script,  
         lemma != "") |>  
  pull(lemma)


```

#### Term frequencies - the script

```{r}

# get word frequencies for the cleaned script tokens
script_word_freq <- table(clean_tokens_azkaban_script)

# to percentage
script_word_freq_percent <- prop.table(script_word_freq) * 100

# show top 50 most frequent words with percentages
sort(script_word_freq_percent, decreasing = TRUE) |> head(50)

```


### Comparison - book vs script

While Ron is mentioned more than Hermione in the book (~750 vs. ~500 times), he is mentioned less than her in the script (~230-240 vs. ~250 times). Similarly, Lupin has more mentions than Hagrid in the book (~400 vs. ~250), but in the script, their mentions are closer, with Lupin still ahead.

Snape has around 250 mentions in the book, while Dumbledore has around 200 mentions. So, in the book, Snape is mentioned more frequently than Dumbledore. However, in the script, their mentions are closer, with Dumbledore at around 150 and Snape at around 200, meaning Dumbledore is mentioned a little less than Snape in the script.

Sirius, despite being the least mentioned in the book (~250), retains a similar position in the script (~150), staying ahead of Dumbledore (~100) in the script but falling behind him in the book.

To summarise:

Amongst the most prominent characters in the Harry Potter Universe, 'Prisoner of the Azkaban' keeps Harry, Ron, and Hermione central to the story while changing the order in favour of Hermione against Ron. Lupin is the main supporting character for both book and script. Although a key character for the book, Hagrid loses his place to be the last in the script amongst all 8 characters, instead, Sirius Black becomes more important. For those who are familiar with the movie, this is an expected result since he had quite a lot of appearances throughout the movie. Snape and Dumbledore maintains their orders, however, Dumbledore is mentioned slightly more in the script compared to his mentions in the book. This also implies that Dumbledore becomes more influential for the movie, however not too much that it would change the order.

```{r}
# to compare main characters
characters <- c("Harry", "Hermione", "Ron", "Sirius", "Snape", "Hagrid", "Dumbledore", "Lupin", "Pettigrew")

#  mentions of each character in both the book and script
book_characters <- clean_tokens_azkaban_book[clean_tokens_azkaban_book %in% characters]
script_characters <- clean_tokens_azkaban_script[clean_tokens_azkaban_script %in% characters]

# frequency tables
book_character_freq <- table(book_characters)
script_character_freq <- table(script_characters)

# combining data for plotting
character_comparison_df <- data.frame(
  character = rep(names(book_character_freq), 2),
  frequency = c(as.vector(book_character_freq), as.vector(script_character_freq)),
  version = rep(c("Book", "Script"), each = length(book_character_freq))
)

# let's plot

ggplot(character_comparison_df, aes(x = reorder(character, frequency), y = frequency, fill = version)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(title = "Character Mentions in Book vs Script", x = "Character", y = "Frequency") +
  scale_fill_manual(values = c("Book" = "#B40000", "Script" = "#0033A0"))  


```



## Sentiment Analysis - book vs script

The book:

Harry - Ron - Hermione - Lupin - Hagrid - Snape - Dubmledore - Sirius - Pettigrew

Negative - positive - fear - anger - sadness - trust - disgust - joy - surprise

The script:
Harry - Hermione - Ron - Lupin - Sirius - Snape - Dumbledore - Hagrid - Pettigrew

Negative - positive - sadness - anger - trust - fear - disgust - joy - surprise

### The book



```{r}
# get sentiment scores for the book
book_sentiment <- get_nrc_sentiment(azkaban_book$text)  # Sentiment scores for the book

# calculate overall sentiment for the book (sum of positive and negative scores)
book_sentiment_score <- book_sentiment %>%
  summarise(across(everything(), sum)) %>%
  unlist()  # Convert the summary to a vector

# create a data frame for the sentiment comparison
sentiment_categories <- c("positive", "negative", "anger", "disgust", "fear", "joy", "sadness", "surprise", "trust")

# Ensure the sentiment score vector has the correct length
book_sentiment_score <- book_sentiment_score[sentiment_categories]  # Align to categories

sentiment_comparison_df <- data.frame(
  sentiment = sentiment_categories,
  book_score = book_sentiment_score
)

# Sort the sentiment data in descending order
sentiment_comparison_df <- sentiment_comparison_df %>%
  arrange(desc(book_score))

# plot sentiment analysis for the book
ggplot(sentiment_comparison_df, aes(x = reorder(sentiment, book_score), y = book_score, fill = "Book")) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Sentiment Analysis for the Book", x = "Sentiment", y = "Score") +
  scale_fill_manual(values = c("Book" = "#B40000")) +
  theme_minimal() + 
  coord_flip()


```


### The script



```{r}
# get sentiment scores for the script
script_sentiment <- get_nrc_sentiment(azkaban_script_lines)  # Sentiment scores for the script

# calculate overall sentiment for the script (sum of positive and negative scores)
script_sentiment_score <- script_sentiment %>%
  summarise(across(everything(), sum)) %>%
  unlist()  # Convert the summary to a vector

# create a data frame for the sentiment comparison
script_sentiment_score <- script_sentiment_score[sentiment_categories]  # Align to categories

sentiment_comparison_df_script <- data.frame(
  sentiment = sentiment_categories,
  script_score = script_sentiment_score
)

# Sort the sentiment data in descending order
sentiment_comparison_df_script <- sentiment_comparison_df_script %>%
  arrange(desc(script_score))

# plot sentiment analysis for the script
ggplot(sentiment_comparison_df_script, aes(x = reorder(sentiment, script_score), y = script_score, fill = "Script")) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Sentiment Analysis for the Script", x = "Sentiment", y = "Score") +
  scale_fill_manual(values = c("Script" = "#0033A0")) +
  theme_minimal() + 
  coord_flip()

```


























